Titeln på avhandlingens språk

### 2023-04-03 body/part1/chapter1-introduction/background.tex
wooork

\subsubsection{\glsfmtlong{WCA}}
\glsreset{WCA}

@@

\todo[inline]{
    - Section 1.1.4: Goes again in the right direction, but details are missing/off. First of all, NCS is really a research direction, not necessarily a system type (every control system is more or less networked, but networking effects have not been always captured in the theoretical analysis/design of the controller). Second, maybe you stress more the working principle in terms workloads and requirements. In a sense, you can work out a pattern how you structure the discussion in Section 1.1.3, and reuse that pattern to a large extent in 1.1.4, that would be neat (both are closed-loop applications with a lot of similar, but also some different properties and requirements). 
}

\glspl{NCS} are a class of control systems that involve the integration of communication networks into traditional control systems.
A control system is a collection of devices or subsystems that work together to achieve a desired behavior or outcome of a physical system.
These systems are composed of
\begin{inlineenum}
    \item a \emph{plant}
    \item \emph{sensors}
    \item \emph{actuators}
    \item a \emph{controller}
\end{inlineenum}.
Plant refers to the physical system being controlled.
Sensors sample and encode the state of the plant and transmit it to the controller.
The controller processes these inputs, and generates control signals to be sent to the actuators, which modify the plant's behavior.
The overall goal of a control system is to regulate a process or a physical system to meet desired specifications or objectives, such as maintaining a constant temperature, tracking a set trajectory, or regulating the speed of a machine.

\glspl{NCS}~\cite{gupta2010networked} aim to enhance and extend the capabilities of traditional control systems by decoupling the controller from the plant, sensors, and actuators and interconnecting them with general-purpose \acs{TCP}/\acs{IP} networks;
see \cref{fig:csvsncs}.
This allows for remote monitoring and control, which enhances the reliability and safety of the control system, which is particularly useful in applications where real-time monitoring and control are critical, such as in industrial automation and transportation systems.
\glspl{NCS} also enable the integration of multiple control systems, as well as centralized control of multiple plants by a single controller.
This leads to enhanced scalability and flexibility, and allows improved coordination and collaboration among systems.
Furthermore, \glspl{NCS} provide greater access to information, as they enable the exchange of data and information between control components and systems, leading to more informed decision-making and improved performance.

@@

\glspl{NCS} have become increasingly popular in recent years due to advances in communication and control technologies, as well as the increasing demand for real-time, remote, and distributed control.
They have found use in varied applications, such as industrial automation, transportation systems, and smart buildings, and some have attempted to leverage the Cloud for centralized, distributed control in what has been called \emph{Cloud control systems}~\cite{xia2015cloud}
However, depending on the physical system being controlled, \glspl{NCS} can have stringent timing and reliability requirements for communication that conventional Cloud paradigms simply cannot meet~\cite{wan2020efficient}.
This has led their adoption to be limited mostly to industrial environments.
This is about to change, however, as with the advent of Edge Computing, paired with novel wireless communication technologies such as cellular 5G, consumer-grade \glspl{NCS} will be made possible.
These technologies will likely become the backbone of consumer-grade, widely-deployed \glspl{NCS}, enabling real-time capabilities through extremely low end-to-end latencies together with context- and locality-awareness.

>>>>

\subsubsection{\glsfmtlong{WCA}}\label{sec:background:wca}
\glsreset{WCA}

@@

% \todo[inline]{
%     - Section 1.1.4: Goes again in the right direction, but details are missing/off. First of all, NCS is really a research direction, not necessarily a system type (every control system is more or less networked, but networking effects have not been always captured in the theoretical analysis/design of the controller). Second, maybe you stress more the working principle in terms workloads and requirements. In a sense, you can work out a pattern how you structure the discussion in Section 1.1.3, and reuse that pattern to a large extent in 1.1.4, that would be neat (both are closed-loop applications with a lot of similar, but also some different properties and requirements). 
% }

A control system is a collection of devices or subsystems that work together to achieve a desired behavior or outcome of a physical system.
These systems are composed of
\begin{inlineenum}
    \item a \emph{plant}
    \item \emph{sensors}
    \item \emph{actuators}
    \item a \emph{controller}
\end{inlineenum}.
Plant refers to the physical system being controlled.
Sensors sample and encode the state of the plant and transmit it to the controller.
The controller processes these inputs, and generates control signals to be sent to the actuators, which modify the plant's behavior.
The overall goal of a control system is to regulate a process or a physical system to meet desired specifications or objectives, such as maintaining a constant temperature, tracking a set trajectory, or regulating the speed of a machine.

\gls{NCS} research is a research direction in the field of automatic control that focuses on the study and development of control systems that formally account for the effects of incorporating communication networks for data exchange between different components of the system.
Every modern control system is, to some extent, networked in the sense that a communication medium is used to transmit information between components.
The field of \gls{NCS} research formalizes and extends this characteristic by integrating traditional control systems with general-purpose \acs{TCP}/\acs{IP} networks.
\glspl{NCS}~\cite{gupta2010networked} aim to enhance and extend the capabilities of traditional control systems by decoupling the controller from the plant, sensors, and actuators (see \cref{fig:csvsncs}).
This allows for remote monitoring and control, which enhances the reliability and safety of the control system
This is particularly useful in applications where real-time monitoring and control are critical, such as in industrial automation and in transportation systems.
\glspl{NCS} also enable the integration of multiple control systems, as well as centralized control of multiple plants by a single controller.
This leads to enhanced scalability and flexibility, and allows improved coordination and collaboration among systems.
Furthermore, \glspl{NCS} provide greater access to information, as they enable the exchange of data and information between control components and systems, leading to more informed decision-making and improved performance.

@@

\input{tables/ncs-examples}

\glspl{NCS} have become increasingly popular in recent years due to advances in communication and control technologies, as well as the increasing demand for real-time, remote, and distributed control.
They have found use in varied applications, such as industrial automation, transportation systems, and smart buildings, and some have attempted to leverage the Cloud for centralized, distributed control in what has been called \emph{Cloud control systems}~\cite{xia2015cloud}
However, \glspl{NCS} can have timing requirements for communication that conventional Cloud paradigms simply cannot meet~\cite{wan2020efficient}.
Depending on the system being controlled, control systems can have sampling rates ranging from sub-\SI{1}{\hertz} to several \unit{\kilo\hertz}.
Reliability is another concern, as many of the core use cases for \gls{NCS} involve application domains in which failure has the potential to cause bodily harm to humans or significant financial losses.
\cref{tab:controlsystemqos} lists some example use cases for \gls{NCS} and their \gls{QoS} requirements.
These stringent requirements have led their adoption to be limited mostly to industrial environments.
This is about to change, however, as with the advent of \gls{MEC}, consumer-grade \glspl{NCS} will be made possible.
This technology will likely become the backbone of consumer-grade, widely-deployed \glspl{NCS}, enabling real-time capabilities through extremely low end-to-end latencies together with context- and locality-awareness.

### 2023-04-03 body/part1/chapter3-contributions/methodology.tex
wooork

As summarized in \cref{sec:summary_contributions}, this thesis presents a comprehensive study on latency-sensitive applications in Edge Computing.
The primary contribution of this research is a methodology for analyzing and evaluating the performance of these applications.
The methodology is validated through the examination of two initial case studies, \gls{WCA} and \glspl{NCS}.
In addition to the methodology, this thesis includes several other key contributions.
A prototype testbed and accompanying software framework have been developed to provide a platform for conducting experiments in Edge Computing.
Furthermore, a further deeper exploration into the \gls{WCA} case studied yielded the first-ever model of human behavior in \gls{WCA}, which is a significant advancement in the field.
The methodology, the case studies, and the prototype testbed provide valuable insights into the performance of latency-sensitive applications in Edge Computing and can be used as a basis for future research in this area.

@@

\begin{figure}
    \centering
    \includegraphics[width=.8\textwidth]{Figs/methodology}
    \caption{%
        Overview of the proposed methodology for the study of latency-sensitive systems on edge computing infrastructure.
        The  client side of the system is replaced with an emulation executed on a general-purpose computing device.
        The backend/server side of the system along with the network remain unchanged.
    }\label{fig:methodology}
\end{figure}

@@

As discussed in \cref{chap:relatedwork}, most research in this field falls into either simulated or fully-practical real-world approaches.
Our emulated methodology presents several advantages over these alternatives.
To begin with, and perhaps most importantly, it makes our approach portable and applicable to real edge computing deployments.
By limiting the emulation to the workload, our approach can be directly employed on pre-existing edge computing infrastructure in order to benchmark applications before they are deployed.
This makes our approach valuable to not only to the research community, but also to system designers and application developers.

In hand with the previous point, it allows for easier scaling compared to real-world studies.
The systems we target exhibit a centralized nature in which a potentially large number of clients offload computation to a single central compute node.
The client-side component of the applications of interest for this research is often complex to scale.
Consider a system owner interested in studying the performance of some edge computing setup for latency-sensitive applications.
In the case of interactive and immersive applications, scaling the number of human users is a significant challenge, as recruiting participants is an expensive and time-consuming endeavor.
Scaling applications involving \glspl{CPS} such as \glspl{NCS} can be prohibitively complex and expensive depending on the specific hardware involved.
In many cases, these systems are made-to-order, further increasing the cost as economies of scale cannot be employed.
On the other hand, a simulated approach requires modeling not only the application workload, but also the edge computing system and infrastructure.
Due to the intricate nature of contemporary cloud and edge computing setups, which frequently involve multiple tiers of proxies, load balancers, virtualized computing units, and network functions, the modeling of such systems can quickly become an overly complex endeavor.

@@

We first introduce our methodology and its applicability to \gls{WCA} in \cref{paper:olguinmunoz2018demoscaling,paper:olguinmunoz2019edgedroid}.
The former corresponds to an extended abstract paper which discusses a high level overview of our approach;
the latter presents a deeper, more complete discussion about the implementation together with some first experimental results.
As discussed in \cref{chap:introduction,chap:relatedwork}\todo{Was it really here?}, the main challenge to benchmarking \gls{WCA} --- and other ``human-in-the-loop'' applications on the edge --- concerns the involvement of human beings in their operation.
Humans are unreliable, and, perhaps more importantly, hard to scale.

@@

We discuss an initial case study of our methodology in the context of \glspl{NCS} and similar \glspl{CPS} in \cref{paper:olguinmunoz2022cleave}.
In this work, we present \gls{CLEAVE}, a software-based framework for scalable and repeatable benchmarking of edge-native \aclp{NCS}.
The growing adoption of edge computing has made it important to understand the strengths and weaknesses of different edge concepts for the deployment of such \glspl{CPS}.

\gls{CLEAVE} is a fully virtualized benchmarking framework, inspired by our previous work in \cref{paper:olguinmunoz2018demoscaling,paper:olguinmunoz2019edgedroid} on benchmarking human-in-the-loop applications on the edge.
In the case of \gls{CLEAVE} however, we forgo the trace-based approach of EdgeDroid in favor of direct modeling of the target systems.
\gls{CLEAVE} includes a benchmarking framework and software development kit for the development of emulated physical systems and softwarized controllers.
These virtual \glspl{CPS} can then be deployed on real networks for reparametrizable, repeatable, and reproducible benchmarking.

\gls{CLEAVE} is built using Python 3.8, making it highly extensible and able to use existing libraries.
It is compatible with container technologies such as Docker~\cite{docker}, making it suitable for automated deployment, scaling, and benchmarking on industry-standard edge setups using container orchestration solutions.

@@

\medskip
We study the validity of our methodology and the implementation of the software framework through a simple scenario, illustrated in \cref{fig:cleavesetup}.
An inverted pendulum plant controlled by a proportional-differential controller is deployed on an Edge server over a Wi-Fi link, while video-streaming applications run concurrently on the setup.
The inverted pendulum is a popular benchmark in control-system literature, and the setup allows for straightforward reparametrization and a broad range of experiments.
Video analytics, on the other hand, is a main proposed use case for edge computing, and it is likely that edge \gls{NCS} deployments will be deployed in parallel with such applications in the future.
Using Wi-Fi allows for more experimental freedom, as the system can be adapted to study both best-case and worst-case scenarios with low and high latencies, respectively.

@@

Next, we study the interaction between the control systems and the video-streaming applications.
Six control loops are deployed, as uplink \gls{UDP} traffic from four clients is generated to emulate the load generated by Full-HD video streaming.
Based on the baseline results, we execute scenarios with varying plant sampling rates and repeat each scenario \num{30} times.
Representative results from this experimentation are showcased in \cref{fig:cleaveresults}.
While the baselines might lead us to think that higher sampling rates are always better for the stability of control systems, the results from these experiments show that network and resource contention are important factors when designing \glspl{NCS} for deployment on the edge.
As sampling rates increase, a critical point is eventually reached at which the congestion on the network link is such that it severely impacts the quality of control.
We conclude that adaptive sampling might be thus a viable method for optimizing resource utilization in these systems.

The results of the experiments will be useful for optimizing resource utilization in edge-bound networked control systems.
The results of the experiments with video-streaming applications will add to the understanding of the interaction between control systems and networked applications.
Furthermore, the success of our experimentation shows that our methodology, implemented through \gls{CLEAVE}, makes the repeated experimentation and data collection convenient compared to fully physical setups, while at the same time providing enhanced realism with respect to simulated approaches.

@@

After our initial implementation and experimentation with our methodology for \gls{WCA} applications, we realized that creating a testbed for experimental Edge Computing research for would significantly enhance our work.
Establishing such testbed would provide a dedicated and controlled environment that can accelerate future research by eliminating the need for ad-hoc implementations for each new study.
Moreover, having a standardized baseline would make it easier to compare results across different studies.
THe testbed could serve as a platform for researchers to test new ideas and compare them to the results of previous experiments.
This would enable more efficient and accurate assessments of the methodology's performance and facilitate the development of new methods and techniques.
Additionally, by establishing a testbed, we could provide a platform for other researchers within our group to conduct their research and experiments.

These reasons led to the establishment of the \gls{EXPECA} testbed at \pgls{KTH}.
\gls{EXPECA} is a \gls{SSF}-funded infrastructure project targeting the development and provisioning of an edge computing infrastructure for research into novel applications and network architectures.
It consists of a cluster of hardware-reconfigurable \gls{COTS} computing nodes interconnected using managed switches and \glspl{SDR}.
This allows us to quickly, on-the-fly, and in an automated fashion change the characteristics of the cluster and the network, in order to study different Edge- and Cloud-computing deployments and the applications that run on them.

>>>>

As summarized in \cref{chap:introduction}, this thesis presents a comprehensive study on latency-sensitive applications in Edge Computing.
The primary contribution of this research is a methodology for analyzing and evaluating the performance of these applications.
The methodology is validated through the examination of two initial case studies, \gls{WCA} and \glspl{NCS}.
In addition to the methodology, this thesis includes several other key contributions.
A prototype testbed and accompanying software framework have been developed to provide a platform for conducting experiments in Edge Computing.
Furthermore, a further deeper exploration into the \gls{WCA} case studied yielded the first-ever model of human behavior in \gls{WCA}, which is a significant advancement in the field.
The methodology, the case studies, and the prototype testbed provide valuable insights into the performance of latency-sensitive applications in Edge Computing and can be used as a basis for future research in this area.

@@

\begin{figure}
    \centering
    \includegraphics[width=.8\textwidth]{Figs/methodology}
    \caption{%
        Overview of the proposed methodology for the study of latency-sensitive systems on Edge computing infrastructure.
        The  client side of the system is replaced with an emulation executed on a general-purpose computing device.
        The backend/server side of the system along with the network remain unchanged.
    }\label{fig:methodology}
\end{figure}

@@

% As discussed in \cref{chap:relatedwork}, most research in this field falls into either simulated or fully-practical real-world approaches.
Our emulated methodology presents several advantages over alternatives, simulated or fully-practical approaches.
To begin with, and perhaps most importantly, it makes our approach portable and applicable to real Edge computing deployments.
By limiting the emulation to the workload, our approach can be directly employed on pre-existing Edge computing infrastructure in order to benchmark applications before they are deployed.
This makes our approach valuable to not only to the research community, but also to system designers and application developers.

In hand with the previous point, it allows for easier scaling compared to real-world studies.
The systems we target exhibit a centralized nature in which a potentially large number of clients offload computation to a single central compute node.
The client-side component of the applications of interest for this research is often complex to scale.
Consider a system owner interested in studying the performance of some Edge computing setup for latency-sensitive applications.
In the case of interactive and immersive applications, scaling the number of human users is a significant challenge, as recruiting participants is an expensive and time-consuming endeavor.
Scaling applications involving \glspl{CPS} such as \glspl{NCS} can be prohibitively complex and expensive depending on the specific hardware involved.
In many cases, these systems are made-to-order, further increasing the cost as economies of scale cannot be employed.
On the other hand, a simulated approach requires modeling not only the application workload, but also the Edge computing system and infrastructure.
Due to the intricate nature of contemporary cloud and Edge computing setups, which frequently involve multiple tiers of proxies, load balancers, virtualized computing units, and network functions, the modeling of such systems can quickly become an overly complex endeavor.

@@

We first introduce our methodology and its applicability to \gls{WCA} in \cref{paper:olguinmunoz2018demoscaling,paper:olguinmunoz2019edgedroid}.
The former corresponds to an extended abstract paper which discusses a high level overview of our approach;
the latter presents a deeper, more complete discussion about the implementation together with some first experimental results.
As discussed in \cref{chap:introduction,sec:relwork:xremulation}, the main challenge to benchmarking \gls{WCA} --- and other ``human-in-the-loop'' applications on the Edge --- concerns the involvement of human beings in their operation.
Humans are unreliable, and, perhaps more importantly, hard to scale.

@@

We discuss an initial case study of our methodology in the context of \glspl{NCS} and similar \glspl{CPS} in \cref{paper:olguinmunoz2022cleave}.
In this work, we present \gls{CLEAVE}, a software-based framework for scalable and repeatable benchmarking of Edge-native \aclp{NCS}.
The growing adoption of Edge computing has made it important to understand the strengths and weaknesses of different Edge concepts for the deployment of such \glspl{CPS}.

\gls{CLEAVE} is a fully virtualized benchmarking framework, inspired by our previous work in \cref{paper:olguinmunoz2018demoscaling,paper:olguinmunoz2019edgedroid} on benchmarking human-in-the-loop applications on the Edge.
In the case of \gls{CLEAVE} however, we forgo the trace-based approach of EdgeDroid in favor of direct modeling of the target systems.
\gls{CLEAVE} includes a benchmarking framework and software development kit for the development of emulated physical systems and softwarized controllers.
These virtual \glspl{CPS} can then be deployed on real networks for reparametrizable, repeatable, and reproducible benchmarking.

\gls{CLEAVE} is built using Python 3.8, making it highly extensible and able to use existing libraries.
It is compatible with container technologies such as Docker~\cite{docker}, making it suitable for automated deployment, scaling, and benchmarking on industry-standard Edge setups using container orchestration solutions.

@@

\medskip
We study the validity of our methodology and the implementation of the software framework through a simple scenario, illustrated in \cref{fig:cleavesetup}.
An inverted pendulum plant controlled by a proportional-differential controller is deployed on an Edge server over a Wi-Fi link, while video-streaming applications run concurrently on the setup.
The inverted pendulum is a popular benchmark in control-system literature, and the setup allows for straightforward reparametrization and a broad range of experiments.
Video analytics, on the other hand, is a main proposed use case for Edge computing, and it is likely that Edge \gls{NCS} deployments will be deployed in parallel with such applications in the future.
Using Wi-Fi allows for more experimental freedom, as the system can be adapted to study both best-case and worst-case scenarios with low and high latencies, respectively.

@@

Next, we study the interaction between the control systems and the video-streaming applications.
Six control loops are deployed, as uplink \gls{UDP} traffic from four clients is generated to emulate the load generated by Full-HD video streaming.
Based on the baseline results, we execute scenarios with varying plant sampling rates and repeat each scenario \num{30} times.
Representative results from this experimentation are showcased in \cref{fig:cleaveresults}.
While the baselines might lead us to think that higher sampling rates are always better for the stability of control systems, the results from these experiments show that network and resource contention are important factors when designing \glspl{NCS} for deployment on the Edge.
As sampling rates increase, a critical point is eventually reached at which the congestion on the network link is such that it severely impacts the quality of control.
We conclude that adaptive sampling might be thus a viable method for optimizing resource utilization in these systems.

The results of the experiments will be useful for optimizing resource utilization in Edge-bound networked control systems.
The results of the experiments with video-streaming applications will add to the understanding of the interaction between control systems and networked applications.
Furthermore, the success of our experimentation shows that our methodology, implemented through \gls{CLEAVE}, makes the repeated experimentation and data collection convenient compared to fully physical setups, while at the same time providing enhanced realism with respect to simulated approaches.

@@

After our initial implementation and experimentation with our methodology for \gls{WCA} applications, we realized that creating a testbed for experimental Edge Computing research for would significantly enhance our work.
Establishing such testbed would provide a dedicated and controlled environment that can accelerate future research by eliminating the need for ad-hoc implementations for each new study.
Moreover, having a standardized baseline would make it easier to compare results across different studies.
The testbed could serve as a platform for researchers to test new ideas and compare them to the results of previous experiments.
This would enable more efficient and accurate assessments of the methodology's performance and facilitate the development of new methods and techniques.
Additionally, by establishing a testbed, we could provide a platform for other researchers within our group to conduct their research and experiments.

These reasons led to the establishment of the \gls{EXPECA} testbed at \pgls{KTH}.
\gls{EXPECA} is a \gls{SSF}-funded infrastructure project targeting the development and provisioning of an Edge computing infrastructure for research into novel applications and network architectures.
It consists of a cluster of hardware-reconfigurable \gls{COTS} computing nodes interconnected using managed switches and \glspl{SDR}.
This allows us to quickly, on-the-fly, and in an automated fashion change the characteristics of the cluster and the network, in order to study different Edge- and Cloud-computing deployments and the applications that run on them.

### 2023-04-03 body/part1/chapter3-contributions/wca-implications.tex
wooork

An important portion of the present thesis deals with the modeling of human behavior for the benchmarking of \gls{WCA}.
We focus in particular on \emph{step-based} \gls{WCA}, i.e.\ assistance applications which have as their goal the guidance of a user through a task, using a series of instructions.
In this section, we will provide some definitions relating to the operation of these applications.

@@

\glsreset{TTF}%
From this characterization, we can define key variables relating to human behavior with respect to the sampling behavior of these applications.
However, we first note that the combination of the characteristic of seamless interaction with the context of the user described in the previous section and in \cref{fig:wca}, combined with the sampling behavior described above, lead to an interesting consequence for the perception of responsiveness in \gls{WCA}.
The only points in time at which a user can consciously notice and interact with the application is whenever they are \emph{expecting} feedback.
A corollary of this is that these are therefore also the \emph{only} points in time at which users can notice changes in system responsiveness.
If we understand \ensuremath{t_c} as the point in time at which the user finishes executing the instruction for step \ensuremath{M}, then
\begin{inlineenum}
    \item \ensuremath{t_c - t_0} corresponds to the \emph{step execution time}, that is, the time it has taken the user to complete the instruction
    \item \ensuremath{t_n - t_c} is the \emph{wait time}, an interval of time during which the task has been finished, but no sample capturing the new state has been taken yet
    \item finally, the interval \ensuremath{t_{n + 1} - t_c} we call \emph{\gls{TTF}}, and refers to the time the user spends waiting for a new instruction.
\end{inlineenum}
These values are key to understanding human perception of responsiveness in these systems, and therefore also to the optimization of resource consumption in \gls{WCA}.
We will refer to them extensively in \cref{chap:contributions}.

@@

\todo[inline]{Do we need more experiments?}

>>>>

In order to more easily and accurately discuss the implications of our methodology for \gls{WCA}, we first characterize these applications and provide some definitions relating to their operation.
We focus in particular on \emph{step-based} \gls{WCA}, i.e.\ assistance applications which have as their goal the guidance of a user through a task, using a series of instructions.

@@

\glsreset{TTF}%
From this characterization, we can define key variables relating to human behavior with respect to the sampling behavior of these applications.
However, we first note that the combination of the characteristic of seamless interaction with the context of the user described in \cref{sec:background:wca} and in \cref{fig:wca:operation}, combined with the sampling behavior described above, lead to an interesting consequence for the perception of responsiveness in \gls{WCA}.
The only points in time at which a user can consciously notice and interact with the application is whenever they are \emph{expecting} feedback.
A corollary of this is that these are therefore also the \emph{only} points in time at which users can notice changes in system responsiveness.
If we understand \ensuremath{t_c} as the point in time at which the user finishes executing the instruction for step \ensuremath{M}, then
\begin{inlineenum}
    \item \ensuremath{t_c - t_0} corresponds to the \emph{step execution time}, that is, the time it has taken the user to complete the instruction
    \item \ensuremath{t_n - t_c} is the \emph{wait time}, an interval of time during which the task has been finished, but no sample capturing the new state has been taken yet
    \item finally, the interval \ensuremath{t_{n + 1} - t_c} we call \emph{\gls{TTF}}, and refers to the time the user spends waiting for a new instruction.
\end{inlineenum}
These values are key to understanding human perception of responsiveness in these systems, and therefore also to the optimization of resource consumption in \gls{WCA}.
We will refer to them extensively in \cref{chap:contributions}.

@@
